---
title: "Homework1"
author: "YikunHan42"
output: 
  html_document
date: '2022-04-04'
---
# homework1_part1

## Importing libraries

```{r}
library(tidyverse)
library(stringr)
```

## Reading files & Setting variables

```{r}
data <- read.csv("D://Study/DSBI/Task1/test_id_card_no.csv")
sex =c()
Birthday =c()
Valid = c()
Age = c()
Zone_code = c()
```

## 1

### (1)

```{r}
data %>% filter(str_detect(Id_Card_No,pattern = "^22"))
```


### (2)(3)
```{r}
#flag变量，看是否存在不通过校验码
count = 0
#循环处理
for(i in 1:nrow(data)){
#校验码
  newdata<-data[c(i),2]
  a = substring(newdata, 1:18,1:18)

  if( a[18] == 'X'){a[18]=10}
    sum_ = 0
    for(j in 1:17){
    sum_ = sum_ +  as.numeric(a[j]) * (2**(18-j) %% 11)
    }
    a1 = ((sum_ %% 11) + as.numeric(a[18]))%%11- 1
  if(a[18]=="10"){a[18]='X'}
  
  if(a1 != 0){
    Valid = append(Valid,0)
    count = count + 1
  print("效验码不对的：")
  print(paste(a, sep = "",collapse=""))}
  if(a1 == 0){
      Valid = append(Valid,1)
  }
#50岁  
  byear <- c(paste(a[7:10], sep = "",collapse=""))
  bmonth <- c(paste(a[11:12], sep = "",collapse=""))
  bday <- c(paste(a[13:14], sep = "",collapse=""))
  b <- paste(byear,bmonth,bday, sep = "-")
  today <- Sys.Date()
  gtd <- as.Date(b) 
  differencetime = difftime(today, gtd, units="days")
  if(differencetime>(365*38+366*12)){
    print("年龄大于50：")
    print(paste(a, sep = "",collapse=""))}
#增添变量
  Birthday<-append(x=Birthday,as.Date(b))
  Age<-append(Age,2022-as.numeric(byear))
  Zone_code = append(Zone_code,c(paste(a[1:2], sep = "",collapse="")))
  #记录性别
  if (as.numeric(a[17])%%2==1){
    sex = append(sex,"Male")
  }
  else{
    sex = append(sex,'Female')
  }
}

if(count == 0) print("全部通过校验")
```
## 2

### (1)
```{r}
data <- data %>% mutate(Birthday)
data %>% arrange(Birthday) 
```

### (2)
```{r}
sex = as.factor(sex)
data <- data %>% mutate(sex)
data %>% arrange(sex,desc(Birthday)) 
```
## 3
```{r}
Valid = as.logical(Valid)
data <- data %>% mutate(Valid)
data
```

## 4
### (1)
```{r}
data <- data %>% mutate(Age)
data %>% filter(Valid=TRUE) %>% summarise(mean(Age))
data %>% filter(Valid=TRUE) %>% summarise(median(Age))
```

### (2)
```{r}
data %>% filter(Valid=TRUE) %>% summarise(n())
```

### (3)
```{r}
data %>% filter(Valid=TRUE) %>% summarise(any(Age<30)) 
```

## 5

### (1)
```{r}
data  %>% filter(Valid=TRUE) %>% group_by(sex) %>% summarise(n(),mean(Age))
```

### (2)
```{r}
data  %>% filter(Valid=TRUE) %>% group_by(sex) %>% summarise(n50=sum(Age>50),n_per=mean(Age>50)) 
```

## 6

### (1)
```{r}
data  %>% filter(Valid=TRUE)  %>% filter(Age<=65) %>% filter(!str_detect(Id_Card_No,pattern = "^220101")) %>% arrange(desc(Age))
newdata <- data[,c(2,3,4,5)]
write.csv(newdata,file = 'Out_Id_Card_Data.csv')
```

### (2)

```{r}
data = data %>% mutate(Zone_code)
data %>% filter(Valid=TRUE) %>% group_by(Zone_code,sex) %>% summarise(n_old=sum(Age>=60),n_old_per=mean(Age>=60),n_not_old=sum(Age<60),n_not_old_per=mean(Age<60))
```


# homework1_part2

## Importing libraries

```{r}
library(tidyverse)
library(Hmisc)
library(dplyr)
library(plyr)
library(stringr)
library(lubridate)
library(tidyr)
```

## (1) Importing the file 读入数据

```{r}
## Sys.setlocale(category="LC_ALL",locale="en_US.UTF-8")
data <- read_csv('D://Study/DSBI/Task1/pharmacy_data.csv')
data %>% as_tibble() ->data
```

## (2) Summary view 数据集概况
```{r}
summary(data)
```

## (3) Renaming date & change the data type 重命名列名&改数据类型
```{r}
data <- rename(data,c("日期"="销售日"))
data$销售日 <- ymd(data$销售日)

data
```

## (4) Deleting rows of missing data 删除缺失数据行
```{r}
data <- data[complete.cases(data[,1:2]),]
```

## (5) Adding mean 插值填补
```{r}
data <- data %>% group_by(品名)
data$数量[is.na(data$数量)] <- mean(data$数量,na.rm=TRUE)
```

## (6) Excluding rows 数量为负的行排除
```{r}
data <- data %>% filter(数量>0)
```

## (7) Descending order 降序排列
```{r}
data %>% arrange(desc(销售日))
```

## (8) Adding line 增加新列
```{r}
data <- data %>% mutate(折扣率= (金额-折扣金额)/金额)
data
```

## (9) Statistic about sale 统计总实际销售额、总销售单数
```{R}
data %>% summarise(n_sale=sum(折扣金额), n_num=sum(数量))
```

## (10) Statistic by commodity 统计各种商品的指标
```{R}
data %>% group_by(品名) %>% dplyr::summarise(n_num=n(),n_sale=sum(金额),n_average=sum(金额)/sum(数量))
```

## (11) Statistic by month 按月统计
```{r}
data %>% 
  mutate(年=year(销售日)) %>% 
  mutate(月=month(销售日)) %>% 
  mutate(年_月=str_c(年,月,sep = "-")) -> data
data %>% 
  group_by(年_月) %>% 
  dplyr::summarise(
    销售单数 = sum(数量),
    总销售额 = sum(金额),
    平均价格 = mean(金额)) 
```

## (12) Statistic by customer 统计每位顾客每月的实际消费金额
```{R}
data %>% 
  mutate(顾客_月=str_c(顾客编号,年_月,sep = "-"))-> data
data %>% 
  group_by(顾客_月) %>% 
  dplyr::summarise(
    总消费额=sum(折扣金额))
```
